Programming Fundamentals and Python

### 2007-08-30 slides/introduction.tex
/doc/slides/*.tex * updated for 0.9
/nltk/wordnet/browse.py
* accepted Jussi Salmela's fixes


svn/trunk@5325

\begin{frame}[fragile]
  \frametitle{Searching Pronunciation Dictionary}
  \tiny

\begin{verbatim}
    >>> from nltk.corpora import cmudict
    >>> from string import join
    >>> for word, num, pron in cmudict.raw():
    ...     stress_pattern = join(c for c in join(pron) if c in "012")
    ...     if stress_pattern.endswith("1 0 0 0 0"):
    ...         print word, "/", join(pron)
    ACCUMULATIVELY / AH0 K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    AGONIZINGLY / AE1 G AH0 N AY0 Z IH0 NG L IY0
    CARICATURIST / K EH1 R AH0 K AH0 CH ER0 AH0 S T
    CIARAMITARO / CH ER1 AA0 M IY0 T AA0 R OW0
    CUMULATIVELY / K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    DEBENEDICTIS / D EH1 B EH0 N AH0 D IH0 K T AH0 S
    DELEONARDIS / D EH1 L IY0 AH0 N AA0 R D AH0 S
    FORMALIZATION / F AO1 R M AH0 L AH0 Z EY0 SH AH0 N
    GIANNATTASIO / JH AA1 N AA0 T AA0 S IY0 OW0
    HYPERSENSITIVITY / HH AY2 P ER0 S EH1 N S AH0 T IH0 V AH0 T IY0
    IMAGINATIVELY / IH2 M AE1 JH AH0 N AH0 T IH0 V L IY0
    INSTITUTIONALIZES / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z AH0 Z
    INSTITUTIONALIZING / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z IH0 NG
    MANGIARACINA / M AA1 N JH ER0 AA0 CH IY0 N AH0
    SPIRITUALIST / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S
    SPIRITUALLY / S P IH1 R IH0 CH AH0 W AH0 L IY0
    UNALIENABLE / AH0 N EY1 L IY0 EH0 N AH0 B AH0 L
    UNDERKOFFLER / AH1 N D ER0 K AH0 F AH0 L ER0
\end{verbatim}
\end{frame}

\subsection{Minimal Sets from Lexicon}
\begin{frame}[fragile]
  \frametitle{Minimal Sets from Lexicon}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import shoebox
    >>> from nltk.utilities import MinimalSet
    >>> length, position, min = 4, 1, 3
    >>> lexemes = [field[1].lower() for entry in shoebox.raw('rotokas')
    ...		   for field in entry if field[0] == 'lx']
    >>> ms = MinimalSet()
    >>> for lex in lexemes:
    ...     if len(lex) == length:
    ...         context = lex[:position] + '_' + lex[position+1:]
    ...         target = lex[position]
    ...         ms.add(context, target, lex)
    >>> for context in ms.contexts(3):
    ...     for target in ms.targets():
    ...         print "%-4s" % ms.display(context, target, "-"),
    ...     print
    kasi -    kesi kusi kosi
    kava -    -    kuva kova
    karu kiru keru kuru koru
    kapu kipu -    -    kopu
    karo kiro -    -    koro
    kari kiri keri kuri kori
    kapa -    kepa -    kopa
    kara kira kera -    kora
    kaku -    -    kuku koku
    kaki kiki -    -    koki
\end{verbatim}
\end{frame}

\subsection{Modelling Text Genres}
\begin{frame}[fragile]
  \frametitle{Modelling Text Genres}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import genesis
    >>> from nltk.probability import ConditionalFreqDist
    >>> from nltk.utilities import print_string
    >>> cfdist = ConditionalFreqDist()
    >>> prev = None
    >>> for word in genesis.raw():
    ...     word = word.lower()
    ...     cfdist[prev].inc(word)
    ...     prev = word
    >>> words = []
    >>> prev = 'lo,'
    >>> for i in range(99):
    ...     words.append(prev)
    ...     for word in cfdist[prev].sorted_samples():
    ...         if word not in words:
    ...             break
    ...     prev = word
    >>> print_string(join(words))
    lo, it came to the land of his father and he said, i will not be a
    wife unto him, saying, if thou shalt take our money in their kind,
    cattle, in thy seed after these are my son from off any more than all
    that is this day with him into egypt, he, hath taken away unawares to
    pass, when she bare jacob said one night, because they were born two
    hundred years old, as for an altar there, he had made me out at her
    pitcher upon every living creature after thee shall come near her:
    yea,
\end{verbatim}
\end{frame}

\subsection{Exploring Syntax}
\begin{frame}[fragile]
  \frametitle{Exploring Syntax}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import treebank
    >>> from string import join
    >>> def vp_conj(tree):
    ...     if tree.node == 'VP' and len(tree) == 3 and tree[1].leaves() == ['but']:
    ...         return True
    ...     else:
    ...         return False
    >>> for tree in treebank.parsed():
    ...     for vp1,conj,vp2 in tree.subtrees(vp_conj):
    ...         print join(child.node for child in vp1), "*BUT*", join(child.node for child in vp2)
    VBP ADVP-TMP PP-PRD PP *BUT* VBP VP
    VBZ VP *BUT* VBZ NP PP-CLR
    PP-TMP VBZ VP *BUT* VBD ADVP-TMP S
    VBZ SBAR *BUT* VBZ SBAR
    VBD SBAR *BUT* VBD RB VP
    VBD SBAR *BUT* VBD S
    VBP NP-PRD *BUT* VBP RB ADVP-TMP VP
    VBN PP PP-TMP *BUT* ADVP-TMP VBN NP
    MD VP *BUT* VBZ NP SBAR-ADV
    VBD ADVP-CLR *BUT* VBD NP
    VBN NP PP *BUT* VBN NP PP SBAR-PRP
    VBD NP *BUT* MD RB VP
    VBD NP PP-CLR *BUT* VBD PRT NP
    VBZ S *BUT* MD VP
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Installing Python and NLTK}
\begin{enumerate}
\item Install Python, Install NLTK, NLTK-Corpora
\item Optional: Numeric, Matplotlib
\item Set environment variable \verb|NLTK_CORPORA|
\end{enumerate}

@@

\begin{itemize}
\item  http://nltk.sourceforge.net/install.html
\item  CDROM: /install.html
\end{itemize}

>>>>

\begin{frame}[fragile]
  \frametitle{Searching Pronunciation Dictionary}
  \scriptsize

\begin{verbatim}
ACCUMULATIVELY / AH0 K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
AGONIZINGLY / AE1 G AH0 N AY0 Z IH0 NG L IY0
CARICATURIST / K EH1 R AH0 K AH0 CH ER0 AH0 S T
CIARAMITARO / CH ER1 AA0 M IY0 T AA0 R OW0
CUMULATIVELY / K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
DEBENEDICTIS / D EH1 B EH0 N AH0 D IH0 K T AH0 S
DELEONARDIS / D EH1 L IY0 AH0 N AA0 R D AH0 S
FORMALIZATION / F AO1 R M AH0 L AH0 Z EY0 SH AH0 N
GIANNATTASIO / JH AA1 N AA0 T AA0 S IY0 OW0
HYPERSENSITIVITY / HH AY2 P ER0 S EH1 N S AH0 T IH0 V AH0 T IY0
IMAGINATIVELY / IH2 M AE1 JH AH0 N AH0 T IH0 V L IY0
INSTITUTIONALIZES / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z AH0 Z
INSTITUTIONALIZING / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z IH0 NG
MANGIARACINA / M AA1 N JH ER0 AA0 CH IY0 N AH0
SPIRITUALIST / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T
SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T S
SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S S
SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S
SPIRITUALLY / S P IH1 R IH0 CH AH0 W AH0 L IY0
UNALIENABLE / AH0 N EY1 L IY0 EH0 N AH0 B AH0 L
UNDERKOFFLER / AH1 N D ER0 K AH0 F AH0 L ER0
\end{verbatim}
\end{frame}

\subsection{Minimal Sets from Lexicon}
\begin{frame}[fragile]
  \frametitle{Minimal Sets from Lexicon}
  \small
\begin{verbatim}
kasi -    kesi kusi kosi
kava -    -    kuva kova
karu kiru keru kuru koru
kapu kipu -    -    kopu
karo kiro -    -    koro
kari kiri keri kuri kori
kapa -    kepa -    kopa
kara kira kera -    kora
kaku -    -    kuku koku
kaki kiki -    -    koki
\end{verbatim}
\end{frame}

\subsection{Modelling Text Genres}
\begin{frame}[fragile]
  \frametitle{Modelling Text Genres}
  \small
\begin{verbatim}
lo, it came to the land of his father and he said, i will not be a
wife unto him, saying, if thou shalt take our money in their kind,
cattle, in thy seed after these are my son from off any more than all
that is this day with him into egypt, he, hath taken away unawares to
pass, when she bare jacob said one night, because they were born two
hundred years old, as for an altar there, he had made me out at her
pitcher upon every living creature after thee shall come near her:
yea,
\end{verbatim}
\end{frame}

\subsection{Exploring Syntax}
\begin{frame}[fragile]
  \frametitle{Exploring Syntax}
  \scriptsize
\begin{verbatim}
VBP ADVP-TMP PP-PRD PP *BUT* VBP VP
VBZ VP *BUT* VBZ NP PP-CLR
PP-TMP VBZ VP *BUT* VBD ADVP-TMP S
VBZ SBAR *BUT* VBZ SBAR
VBD SBAR *BUT* VBD RB VP
VBD SBAR *BUT* VBD S
VBP NP-PRD *BUT* VBP RB ADVP-TMP VP
VBN PP PP-TMP *BUT* ADVP-TMP VBN NP
MD VP *BUT* VBZ NP SBAR-ADV
VBD ADVP-CLR *BUT* VBD NP
VBN NP PP *BUT* VBN NP PP SBAR-PRP
VBD NP *BUT* MD RB VP
VBD NP PP-CLR *BUT* VBD PRT NP
VBZ S *BUT* MD VP
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Installing Python and NLTK}
\begin{enumerate}
\item Install Python, Install NLTK, NLTK-Data
\item Optional: Numpy, Matplotlib
\item Set environment variable \verb|NLTK_DATA|
\end{enumerate}

@@

\begin{itemize}
\item  http://nltk.org/index.php/Installation
\item  CDROM: /install.html
\end{itemize}

### 2007-08-30 slides/programming.tex
fixes to get slides to build
svn/trunk@5031
/doc/slides/*.tex * updated for 0.9
/nltk/wordnet/browse.py
* accepted Jussi Salmela's fixes


svn/trunk@5325

\begin{verbatim}
  >>> from nltk.corpora import gutenberg
  >>> count = {}
  >>> for word in gutenberg.raw('shakespeare-macbeth'):
  ...     word = word.lower()
  ...     if word not in count:
  ...         count[word] = 0
  ...     count[word] += 1
\end{verbatim}

@@

\begin{verbatim}
  >>> import re
  >>> from nltk.utilities import re_show
  >>> sent = "colourless green ideas sleep furiously"
\end{verbatim}

@@

\begin{verbatim}
  >>> re_show('l', sent)
  co{l}our{l}ess green ideas s{l}eep furious{l}y
  >>> re_show('green', sent)
  colourless {green} ideas sleep furiously
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> re_show('(green|sleep)', sent)
  colourless {green} ideas {sleep} furiously
  >>> re.findall('(green|sleep)', sent)
  ['green', 'sleep']
\end{verbatim}

@@

\begin{verbatim}
  >>> re_show('[^aeiou][aeiou]', sent)
  {co}{lo}ur{le}ss g{re}en{ i}{de}as s{le}ep {fu}{ri}ously
  >>> re.findall('[^aeiou][aeiou]', sent)
  ['co', 'lo', 'le', 're', ' i', 'de', 'le', 'fu', 'ri']
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> from urllib import urlopen
  >>> page = urlopen("http://news.bbc.co.uk/").read()
  >>> page = re.sub('<[^>]*>', '', page)
  >>> page = re.sub('\s+', ' ', page)
  >>> print page[:60]
   BBC NEWS | News Front Page News Sport Weather World Service
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item modules: classes, functions
\item data structures, algorithms
\item importing, e.g. \texttt{from nltk\_lite.utilities import *}
\item conflicts
\item explicit naming: \texttt{from nltk\_lite.utilities import re\_show}
\item higher level module import:

@@

\begin{verbatim}
  >>> from nltk.corpora import gutenberg
  >>> gutenberg.items
  ['austen-emma', 'austen-persuasion', 'austen-sense', 'bible-kjv', 'blake-poems', 'blake-songs', 'chesterton-ball', 'chesterton-brown', 'chesterton-thursday', 'milton-paradise', 'shakespeare-caesar', 'shakespeare-hamlet', 'shakespeare-macbeth', 'whitman-leaves']
  >>> count = 0
  >>> for word in gutenberg.raw('whitman-leaves'):
  ...     count += 1
  >>> print count
  154873
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk.corpora import brown
  >>> brown.items
  ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'j', 'k', 'l', 'm', 'n', 'p', 'r']
  >>> from nltk.corpora import extract
  >>> print extract(0, brown.raw())
  ['The', 'Fulton', 'County', 'Grand', 'Jury', 'said', 'Friday', 'an', 'investigation', 'of', "Atlanta's", 'recent', 'primary', 'election', 'produced', '``', 'no', 'evidence', "''", 'that', 'any', 'irregularities', 'took', 'place', '.']
  >>> print extract(0, brown.tagged())
  [('The', 'at'), ('Fulton', 'np-tl'), ('County', 'nn-tl'), ('Grand', 'jj-tl'), ('Jury', 'nn-tl'), ('said', 'vbd'), ('Friday', 'nr'), ('an', 'at'), ('investigation', 'nn'), ('of', 'in'), ("Atlanta's", 'np$'), ('recent', 'jj'), ('primary', 'nn'), ('election', 'nn'), ('produced', 'vbd'), ('``', '``'), ('no', 'at'), ('evidence', 'nn'), ("''", "''"), ('that', 'cs'), ('any', 'dti'), ('irregularities', 'nns'), ('took', 'vbd'), ('place', 'nn'), ('.', '.')]
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk.corpora import treebank
  >>> print extract(0, treebank.parsed())
  (S:
    (NP-SBJ:
      (NP: (NNP: 'Pierre') (NNP: 'Vinken'))
      (,: ',')
      (ADJP: (NP: (CD: '61') (NNS: 'years')) (JJ: 'old'))
      (,: ','))
    (VP:
      (MD: 'will')
      (VP:
        (VB: 'join')
        (NP: (DT: 'the') (NN: 'board'))
        (PP-CLR:
          (IN: 'as')
          (NP: (DT: 'a') (JJ: 'nonexecutive') (NN: 'director')))
        (NP-TMP: (NNP: 'Nov.') (CD: '29'))))
    (.: '.'))
\end{verbatim}
\end{frame}

>>>>

\begin{verbatim}
  >>> import nltk
  >>> count = {}
  >>> for word in nltk.corpus.gutenberg.words('shakespeare-macbeth'):
  ...     word = word.lower()
  ...     if word not in count:
  ...         count[word] = 0
  ...     count[word] += 1
\end{verbatim}

@@

\begin{verbatim}
  >>> import nltk, re
  >>> sent = "colourless green ideas sleep furiously"
\end{verbatim}

@@

\begin{verbatim}
  >>> nltk.re_show('l', sent)
  co{l}our{l}ess green ideas s{l}eep furious{l}y
  >>> nltk.re_show('green', sent)
  colourless {green} ideas sleep furiously
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> nltk.re_show('(green|sleep)', sent)
  colourless {green} ideas {sleep} furiously
  >>> re.findall('(green|sleep)', sent)
  ['green', 'sleep']
\end{verbatim}

@@

\begin{verbatim}
  >>> nltk.re_show('[^aeiou][aeiou]', sent)
  {co}{lo}ur{le}ss g{re}en{ i}{de}as s{le}ep {fu}{ri}ously
  >>> re.findall('[^aeiou][aeiou]', sent)
  ['co', 'lo', 'le', 're', ' i', 'de', 'le', 'fu', 'ri']
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> from urllib import urlopen
  >>> page = urlopen("http://news.bbc.co.uk/").read()
  >>> text = nltk.clean_html(page)
  >>> print text[:60]
   BBC NEWS | News Front Page News Sport Weather World Service
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item modules: classes, functions
\item data structures, algorithms
\item importing, e.g. \texttt{import nltk}

@@

\begin{verbatim}
  >>> nltk.corpus.gutenberg.items
  ['austen-emma', 'austen-persuasion', 'austen-sense', 'bible-kjv', 'blake-poems', 'blake-songs', 'chesterton-ball', 'chesterton-brown', 'chesterton-thursday', 'milton-paradise', 'shakespeare-caesar', 'shakespeare-hamlet', 'shakespeare-macbeth', 'whitman-leaves']
  >>> count = 0
  >>> for word in nltk.corpus.gutenberg.words('whitman-leaves'):
  ...     count += 1
  >>> print count
  154873
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> nltk.corpus.brown.items
  ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'j', 'k', 'l', 'm', 'n', 'p', 'r']
  >>> print nltk.corpus.brown.words('a')
  ['The', 'Fulton', 'County', 'Grand', 'Jury', 'said', 'Friday', 'an', 'investigation', 'of', "Atlanta's", 'recent', 'primary', 'election', 'produced', '``', 'no', 'evidence', "''", 'that', 'any', 'irregularities', 'took', 'place', '.']
  >>> print nltk.corpus.brown.tagged_sents('a')
  [('The', 'at'), ('Fulton', 'np-tl'), ('County', 'nn-tl'), ('Grand', 'jj-tl'), ('Jury', 'nn-tl'), ('said', 'vbd'), ('Friday', 'nr'), ('an', 'at'), ('investigation', 'nn'), ('of', 'in'), ("Atlanta's", 'np$'), ('recent', 'jj'), ('primary', 'nn'), ('election', 'nn'), ('produced', 'vbd'), ('``', '``'), ('no', 'at'), ('evidence', 'nn'), ("''", "''"), ('that', 'cs'), ('any', 'dti'), ('irregularities', 'nns'), ('took', 'vbd'), ('place', 'nn'), ('.', '.')]
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> print nltk.corpus.treebank.parsed_sents('wsj_0001')[0]
  (S:
    (NP-SBJ:
      (NP: (NNP: 'Pierre') (NNP: 'Vinken'))
      (,: ',')
      (ADJP: (NP: (CD: '61') (NNS: 'years')) (JJ: 'old'))
      (,: ','))
    (VP:
      (MD: 'will')
      (VP:
        (VB: 'join')
        (NP: (DT: 'the') (NN: 'board'))
        (PP-CLR:
          (IN: 'as')
          (NP: (DT: 'a') (JJ: 'nonexecutive') (NN: 'director')))
        (NP-TMP: (NNP: 'Nov.') (CD: '29'))))
    (.: '.'))
\end{verbatim}
\end{frame}

% \section{Miscellaneous Python Features}

% \subsection{Character Encoding}

% \begin{frame}
% \begin{itemize}

% \end{itemize}
% \end{frame}
